\section[Uniqueness and Canonical Status]{Uniqueness and Canonical Status of the Symmetry-Induced Complex Structure}

In the preceding section, we constructed a complex structure operator
\[
J := A(-A^2)^{-1/2}
\]
on the invariant subspace
\[
\mathcal{K}_1 = \overline{\mathrm{ran}(A)} = \ker(A)^\perp.
\]

We now analyze the extent to which this complex structure is uniquely determined by the symmetry generator.

\subsection{Symmetry-compatible complex structures}

We begin by formalizing compatibility with the symmetry.

\begin{definition}
A symmetry-compatible complex structure on \(\mathcal{K}_1\) is a bounded linear operator
\[
J' : \mathcal{K}_1 \to \mathcal{K}_1
\]
satisfying:

\begin{enumerate}

\item Complex structure condition:
\[
(J')^2 = -I,
\]

\item Inner product preservation:
\[
\langle J'\psi, J'\phi \rangle
=
\langle \psi, \phi \rangle
\quad \forall \psi,\phi \in \mathcal{K}_1,
\]

\item Symmetry compatibility (domain-qualified commutation):
\[
J'(\mathcal{D}(A)) \subset \mathcal{D}(A),
\qquad
J'A\psi = AJ'\psi
\quad \forall \psi \in \mathcal{D}(A).
\]

\end{enumerate}

\end{definition}

This condition is equivalent to commutation with the strongly continuous symmetry group:
\[
J'T(t) = T(t)J'
\quad \forall t \in \mathbb{R},
\]
by Stone's theorem for strongly continuous orthogonal groups
\cite[Theorem VIII.4]{ReedSimonI}.

Thus symmetry-compatible complex structures lie in the commutant of the symmetry representation.

\subsection{Spectral decomposition and multiplicity structure}

Since \(A\) is skew-adjoint, the operator \(-A^2\) is self-adjoint and positive semidefinite. By the spectral theorem, there exists a projection-valued measure \(E(\cdot)\) such that
\[
-A^2 = \int_0^\infty \lambda \, dE(\lambda),
\]
and the Hilbert space decomposes into spectral subspaces
\[
\overline{\mathcal V}
=
E(\{0\})\overline{\mathcal V}
\;\oplus\;
E((0,\infty))\overline{\mathcal V}
=
\mathcal K_0 \oplus \mathcal K_1
\cite[Theorem VII.2]{ReedSimonI}.
\]

In a spectral representation of \(-A^2\), the invariant subspace \(\mathcal{K}_1\) may be realized as a direct integral
\[
\mathcal{K}_1
\cong
\int_{(0,\infty)}^\oplus H_\lambda \, d\mu(\lambda),
\]
where each fiber \(H_\lambda\) is a Hilbert space whose dimension equals the spectral multiplicity at \(\lambda\)
\cite[Section VII.2]{ReedSimonI}.

In this representation, the operator \(-A^2\) acts as multiplication:
\[
((-A^2)\psi)(\lambda) = \lambda \psi(\lambda),
\]
and operators commuting with the symmetry act fiberwise.

Multiplicity-one fibers correspond to irreducible symmetry sectors, while higher multiplicity fibers admit additional commuting operators acting internally within multiplicity spaces.

\subsection{Commutant structure on irreducible sectors}

We now characterize symmetry-compatible operators on irreducible sectors.

\begin{definition}
A symmetry-invariant subspace \(\mathcal S \subset \mathcal K_1\) is irreducible if it contains no nontrivial closed subspace invariant under the symmetry group \(T(t)\).
\end{definition}

On such a subspace, the generator restricts to a skew-adjoint operator with spectrum consisting of a single positive frequency \(\omega>0\). The symmetry acts as planar rotation:
\[
T(t)|_{\mathcal S}
=
\begin{pmatrix}
\cos(\omega t) & -\sin(\omega t) \\
\sin(\omega t) & \cos(\omega t)
\end{pmatrix}
\]
in a suitable orthonormal basis
\cite{HallQM,VaradarajanGeometryQM}.

For real irreducible rotation representations, the commutant algebra consists precisely of operators of the form
\[
aI + bJ,
\quad a,b \in \mathbb R,
\]
where \(J\) is the canonical complex structure operator obtained via functional calculus of the generator.

This follows from the classification of irreducible real representations of one-parameter rotation groups and their commutant algebras
\cite{KadisonRingrose,VaradarajanGeometryQM}.

\subsection{Sectorwise uniqueness}

We now establish uniqueness of symmetry-compatible complex structure on irreducible sectors.

\begin{theorem}[Sectorwise uniqueness up to sign]
Let \(\mathcal S \subset \mathcal K_1\) be an irreducible symmetry-invariant subspace.

Then any symmetry-compatible complex structure operator
\[
J' : \mathcal S \to \mathcal S
\]
satisfies
\[
J' = \pm J
\quad \text{on } \mathcal S.
\]
\end{theorem}

\begin{proof}

Since \(J'\) commutes with the symmetry and \(\mathcal S\) is irreducible, the commutant characterization implies that
\[
J' = aI + bJ
\]
for some real scalars \(a,b\).

The complex structure condition gives
\[
(J')^2 = -I.
\]

Substituting,
\[
(aI + bJ)^2
=
(a^2 - b^2)I + 2abJ.
\]

Equating with \(-I\) yields
\[
2ab = 0,
\qquad
a^2 - b^2 = -1.
\]

If \(b=0\), then \(J' = aI\), which cannot satisfy \((J')^2 = -I\) for real \(a\).

Thus \(a=0\), and
\[
J' = bJ.
\]

The remaining condition gives
\[
b^2 = 1,
\]
and therefore
\[
J' = \pm J.
\]

\end{proof}

\subsection{Multiplicity and global structure}

When spectral multiplicity exceeds one, the commutant algebra includes additional operators acting within multiplicity fibers
\cite{KadisonRingrose,Dixmier}.

Such operators correspond to internal symmetries of multiplicity spaces and are not uniquely determined by the symmetry generator itself.

By contrast, the operator
\[
J = A(-A^2)^{-1/2}
\]
is defined canonically through functional calculus of the generator and requires no additional choices.

It therefore provides a distinguished symmetry-compatible complex structure determined directly by the symmetry representation.

\subsection{Sign ambiguity and canonical equivalence}

The operator \(-J\) also satisfies all defining properties of a symmetry-compatible complex structure.

This corresponds to complex conjugation and does not alter the isomorphism class of the resulting complex Hilbert space.

Thus, on each irreducible symmetry sector, the symmetry determines a unique compatible complex structure up to sign.

Globally, the functional-calculus operator \(J\) provides the canonical symmetry-induced complex structure associated with the one-parameter symmetry generator.
